---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data for this hands-on session comes from a published RNA-seq experiment where airway smooth muscle cells were treated with dexamethasone, a synthetic glucocorticoid steroid with anti-inflammatory effects (Himes et al. 2014).

Glucocorticoids are used, for example, by people with asthma to reduce inflammation of the airways. The anti-inflammatory effects on airway smooth muscle (ASM) cells has been known for some time but the underlying molecular mechanisms are unclear.


```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)

metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

How many genes are we dealing with?
```{r}
nrow(counts)
```


We want to calculate the average counts per gene for control samples and for the drug treated samples. We will then compare these values to see if there is a difference overall between control and treated samples for a given gene.


This is our `metadata` info that tells me where our control and treated samples are (i.e. which columns in `counts`)
```{r}
metadata
```

Now I want access to the control only rows in the metadata
```{r}
# metadata$id[metadata$dex == "control"] is one way to find it
control <- metadata[metadata$dex == "control",]
```

Now use the $id of the control data to access the control columns in `counts` and calculate the average. `nrow()` function tells use number of samples in control. 
```{r}
control.mean <- rowSums(counts[ ,control$id]) / nrow(control)
names(control.mean) <- counts$ensgene
```

Now lets do the same for *treated* samples
```{r}
treated <- metadata[metadata$dex == "treated", ]
treated.mean <- rowSums(counts[ , treated$id])/nrow(treated)
```
We will combine our meancount data for bookeeping purposes. 
```{r}
meancounts <- data.frame(control.mean, treated.mean)
meancounts
```

```{r}
plot(meancounts)
```

Lets make this plot on a log scale to keep dots from bunching
```{r}
plot(meancounts, log= "xy")
```
