---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data for this hands-on session comes from a published RNA-seq experiment where airway smooth muscle cells were treated with dexamethasone, a synthetic glucocorticoid steroid with anti-inflammatory effects (Himes et al. 2014).

Glucocorticoids are used, for example, by people with asthma to reduce inflammation of the airways. The anti-inflammatory effects on airway smooth muscle (ASM) cells has been known for some time but the underlying molecular mechanisms are unclear.


```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)

metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

How many genes are we dealing with?
```{r}
nrow(counts)
```


We want to calculate the average counts per gene for control samples and for the drug treated samples. We will then compare these values to see if there is a difference overall between control and treated samples for a given gene.


This is our `metadata` info that tells me where our control and treated samples are (i.e. which columns in `counts`)
```{r}
metadata
```

Now I want access to the control only rows in the metadata
```{r}
# metadata$id[metadata$dex == "control"] is one way to find it
control <- metadata[metadata$dex == "control",]
```

Now use the $id of the control data to access the control columns in `counts` and calculate the average. `nrow()` function tells use number of samples in control. 
```{r}
control.mean <- rowSums(counts[ ,control$id]) / nrow(control)
names(control.mean) <- counts$ensgene
```

Now lets do the same for *treated* samples
```{r}
treated <- metadata[metadata$dex == "treated", ]
treated.mean <- rowSums(counts[ , treated$id])/nrow(treated)
```
We will combine our meancount data for bookeeping purposes. 
```{r}
meancounts <- data.frame(control.mean, treated.mean)
meancounts
```

```{r}
plot(meancounts)
```

Lets make this plot on a log scale to keep dots from bunching
```{r}
plot(meancounts, log= "xy")
```

Here we calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the head() or the View() function for example.
```{r}
meancounts$log2fc <-
    log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
    head(meancounts)
```

Remove any gene that has zero average counts.
```{r}
#example of which() function and arr.ind
x <- data.frame(c(1,1,10,0), c(0,1,10,0))
x == 0 
which(x==0, arr.ind = T)
ind <- unique(which(x==0, arr.ind = T)[, "row"])
x[ind,]
```

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

How many genes do I have left?
```{r}
nrow(mycounts)
```
A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2. Let’s filter the dataset both ways to see how many genes are up or down-regulated.

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
sum(down.ind)
```

##Do this properly with DESeq2
```{r}
library(DESeq2)
citation("DESeq2")
```

Setup inputs for DESeq by specifying countdata, metadata, and the design(i.e. which comparisons we are interested in)
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
res <- results(dds)
```

```{r}
plot(res$log2FoldChange, res$padj)
```
```{r}
plot(res$log2FoldChange, log(res$padj))
```

Let's plot this upright like everyone else does
```{r}
plot(res$log2FoldChange, -log(res$padj))
```

```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
redchecker <- abs(res$log2FoldChange) > 2 
mycols[redchecker]  <- "red" 

bluechecker <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[bluechecker] <- "blue"

#Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```


