---
title: "Class 13: Structure based drug discovery"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Download and clean HIV-Pr receptor structure

Here we will download the 1HSG PDB structure from the online databse, remove the bound drug from Merk and write out a protein only and ligand only new set of PDB files.

We will use the functions bio3d package functions `get.pdb()`, `read.pdb()`, `atom.select()`, and finally `write.pdb()`

```{r}
library(bio3d)

#Download PDB file
file.name <- get.pdb("1hsg")
hiv <- read.pdb(file.name)
```

```{r}
hiv

#select protein only
atom.select(hiv, "protein", value = T )

#select ligand only
atom.select(hiv, "ligand", value = T)
```

Q1: What is the name of the two non protein resid values in this structure? HOH(Water) and MK1

What does resid correspond to and how would you get a listing of all reside values in this structure? resid corresponds to the residue positions. `hiv$atom$resid` works

```{r}
#alternative to atom.select() is trim.pdb()
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")
```


```{r}
#check that ligand is there
library(bio3d.view)
view(lig)
```

