---
title: 'Class 6: R Functions'
author: "Kendall Lin"
date: "1/23/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Functions
## Level 2 heading
### Level 3 heading

read.table("Your File Name", Header = True, sep = '\t')

```{r}
plot(1:5, typ = "l", col = "blue")
```

Lets see more about file import(i.e. reading files into R). The main read function in base R is `read.table()` 

```{r}
t1 <- read.table("test1.txt")
```

We need to add arguments to get this file imported

```{r}
t1<- read.table("test1.txt", header = T, sep = ",")
```

Or we could just use `read.csv()` which as the arguments I want in this case!

```{r}
t1<- read.csv("test1.txt")
```

```{r}
t2<- read.table("test2.txt", sep= "$", header = T)
```

```{r}
t3 <- read.table("test3.txt")
```

#Back to functions
Out first example function:
```{r}
add <- function(x, y = 1){
  #Sum the input x and y
  x+y
}
```

Works with vectors too

```{r}
add(c(1,2,3),c(1,2,9))
```

What is this `range()` function?
```{r}
x<- c(1,3,4,11)
range(x)
```


```{r}
rescale <- function(x) {
 rng <-range(x, na.rm = T)
 (x - rng[1]) / (rng[2] - rng[1])
}
rescale(c(1,2,NA,3, 10))
```

```{r}
rescale3<- function(x, na.rm=TRUE, plot=FALSE) {
  
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 
 print("is it me you are looking for?")
 
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}
```
```{r}
rescale3(x, plot = T)
```

```{r}
library(bio3d)
```

Hands on Section B.
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```
```{r}
class(s1)
str(s1)
```


```{r}
s1$seqres
```


```{r}
aa321(s1$seqres)
```

```{r}
s1$atom
```



```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, typ = "l", col = "blue", lwd= 1)
points(s3.b, typ = "l", col = "red", lwd = 1)
```

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```


Q1. What type of object is returned from the read.pdb() function?
**returns a large list from protein database. Can tell it is a list by using str() function on s1, s2, or s3. It is a list of 8 things and of class "pdb"**

Q2. What does the trim.pdb() function do?
**Produces a new smaller PDB object, containing a subset of atoms, from a given larger PDB object**

Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?
**You can remove the elety="CA" to remove the black and grey rectangles. They represent a character vector of atom names**

Q4. What would be a better plot to compare across the different proteins?
**overlay the graphs using points()**

Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this? HINT: try the rbind(), dist() and hclust() functions together with a resulting dendrogram plot. Look up the documentation to see what each of these functions does.
**s2 and s3 are closer to each other according to hclust(), rbind(), dist() function **

Q6.  How would you generalize the original code above to work with any set of input protein structures?


